---
title: "Merge"
author: "Wesley Burr"
date: "06/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Sheets

```{r}
library("readxl")
```

```{r morgue, warning = FALSE, message = FALSE}
sheets <- excel_sheets("./all_Morgue_cleaned2.xlsx") 
morgue <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./all_Morgue_cleaned2.xlsx",
                     sheet = x) })
names(morgue) <- sheets
```

```{r rest, warning = FALSE, message = FALSE}
sheets <- excel_sheets("./all_REST_cleaned2.xlsx")  # note the diff location
rest <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "./all_REST_cleaned2.xlsx",
                     sheet = x) })
names(rest) <- sheets
```

## Merge to Prepare for PCA

The goal here is to extract all unique compounds from all objects, merge them,
make a reference list, and then create a very large data.frame with rows
being compounds, and columns being the various experimental samples, however
many there are.

```{r}
morgue_compounds <- unique(unlist(lapply(morgue, FUN = function(x) { x$Name })))
rest_compounds <- unique(unlist(lapply(rest, FUN = function(x) { x$Name })))
```

```{r}
all_compounds <- unique(c(morgue_compounds,
                          rest_compounds))
```

```{r}
n_total <- length(morgue) + length(rest)
```

### Create the Data Frame

```{r}
all_data <- data.frame(matrix(data = 0.0, nrow = length(all_compounds),
                              ncol = n_total + 1))
all_data[, 1] <- all_compounds
names(all_data) <- c("Name", names(morgue), names(rest))
```

```{r do_morgue}
for(j in 1:length(morgue)) {
  x <- morgue[[j]]
  x_names <- x$Name
  x_area <- apply(x, MAR = 1, FUN = function(y) { 
    z <- y[seq(3, length(y), 2)]; 
    z <- as.numeric(z);
    mean(z) })
  for(k in 1:length(x_area)) {
    all_data[all_data$Name == x_names[k], j+1] <- x_area[k]
  }
}
```

```{r do_rest}
for(j in 1:length(rest)) {
  x <- rest[[j]]
  x_names <- x$Name
  x_area <- apply(x, MAR = 1, FUN = function(y) { 
    z <- y[seq(3, length(y), 2)]; 
    z <- as.numeric(z);
    mean(z) })
  for(k in 1:length(x_area)) {
    all_data[all_data$Name == x_names[k], j+1+length(morgue)] <- x_area[k]
  }
}
```

```{r}
write.csv(file = "pca_ready_all2.csv", all_data, row.names = FALSE)
```

## Create a 'Mean Centered' version

The idea here is to subtract the mean from the PCA columns, so that the mean
of each column is 0. This can sometimes ... do something. I guess.

```{r}
all_data2 <- data.frame(all_data$Name,
                        apply(all_data[, -1], MAR = 2, FUN = function(x) {
                          x - mean(x)
                        }))
write.csv(file = "pca_ready_all2_meanCenter.csv", all_data2, row.names = FALSE)
```

