---
title: "Extract Files"
author: "Wesley Burr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Working from the Aligned Data instead of Excel Files

Libraries first:

```{r libraries, error = FALSE, message = FALSE, warning = FALSE}
library("readxl")
library("stringr")
library("tidyverse")
```

and then we load the aligned data:

```{r}
load("controls_aligned.rda")
load("samples_aligned_100_50_mc28.rda")
```

## Extract and Save Out the Aligned Files for Morgue

Previously, we were loading one Excel file at a time, with its specific
sheets all loaded into a list. These were called M_DONOR_1 through 
M_DONOR_9, and were the morgue files only. We can replicate this:
they are the objects in the control/sample list that are labeled as MR. They
will have H1 through H9 to refer to the 1 to 9. 

Unfortunately the names aren't accurate in the spreadsheets, so we have to
just do this sequentially. H1 is ok (entries 1:3), H2 is ok (4:6), but
H3 isn't (7:9). The first 24 entries are the records, though - 8*3 = 24, 
and this is samples 1:4, 6:9. 

```{r}
names_obj <- c(1:4, 6:9)
names_obj <- paste("M_DONOR_", names_obj, collapse = NULL, sep = "")

for(j in 1:8) {
  dat <- do.call("rbind", samples_aligned[seq((j-1)*3+1, j*3, 1)])
  dat <- rbind(dat, controls_aligned[[j]]) # add the controls in
  names(dat) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "RT1", 
                  "RT2", "Similarity",
                  "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                  "Type", "Conc", "ConcUnits", "Aligned")
 
  assign(x = names_obj[j], dat)
  save(file = paste0("./dat/", names_obj[j], ".rda"), list = names_obj[j])
}
```

## Extract and Save Out the Aligned Files for REST

Previously, we were loading one Excel file at a time, with its specific
sheets all loaded into a list. These were the files RH1 through RH9, 8
in total.

```{r}
names_obj <- c(1:9)
names_obj <- paste("RH", names_obj, collapse = NULL, sep = "")

controls_ss <- substr(names(controls_aligned), 1, 2)
samples_ss <- substr(names(samples_aligned), 1, 2)

# the j=1 case, H1, is special because it only has one control and one triplicate;
# not the repeats of the others
dat <- do.call("rbind", samples_aligned[which(samples_ss == paste0("H", 1))])
dat <- rbind(dat, controls_aligned[[which(controls_ss == paste0("H", 1))]]) 
                  # only one control; need to [[*]] extract
names(dat) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "RT1", 
                "RT2", "Similarity",
                "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                "Aligned")
 
assign(x = names_obj[1], dat)
save(file = paste0("./dat/", names_obj[1], ".rda"), list = names_obj[1])

# the cases of the others are easier; just have to twiddle with 6:9
for(j in c(2:9)[-4]) {
  dat1 <- do.call("rbind", samples_aligned[which(samples_ss == paste0("H", j))])
  dat2 <- do.call("rbind", controls_aligned[which(controls_ss == paste0("H", j))]) 
  dat <- rbind(dat1, dat2)
  names(dat) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "RT1", 
                  "RT2", "Similarity",
                  "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                  "Aligned")
 
  assign(x = names_obj[j], dat)
  save(file = paste0("./dat/", names_obj[j], ".rda"), list = names_obj[j])
}

```
