---
title: "General Function Cleaning (Normalized)"
author: "Wesley Burr"
date: "2022-11-20"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```

## Cleanup, with Normalization Done First: Morgue

In this file, we use the
Bromobenzene-normalized inputs and then clean.

```{r function}
##
#  clean_common: Function to select, filter, clean, filter and merge 
#  compounds using the logic of:
#  * unique to the samples, not the control, for ALL samples; or
#  * in both samples and control, but much stronger in the control
# 
#  Inputs:
#  * dat: data.frame sourced from merging spreadsheets of GCxGC output
#  * sample_names: names of specific samples (e.g., SS5_Foot_1_a)
#  * control_names: names of specific control samples (e.g., SS5_Foot_1_Control_a)
#  * ratio_par: cut-off for the logic of "in both samples and control" - if this
#      is set very large, will eliminate cross-overs.
#
#  Returns:
#  * samples_keep: list of full data.frames for individual replicates, cleaned
#      down to relevant compounds using above logic
##
clean_common <- function(dat,
                         sample_names,
                         control_names,
                         ratio_par = 2.0) {
  
  samples <- vector("list", length = length(sample_names))
  names(samples) <- sample_names
  controls <- vector("list", length = length(control_names))
  names(controls) <- control_names
  
  # Extract specific samples and controls of interest and
  # dump all but the largest Area example of each compound
  for(j in 1:length(sample_names)) {
    samples[[j]] <- dat %>% subset(Sample == sample_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    samples[[j]] <- samples[[j]][!duplicated(samples[[j]]$Name), ]
  } 
  for(j in 1:length(control_names)) {
    controls[[j]] <- dat %>% subset(Sample == control_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    controls[[j]] <- controls[[j]][!duplicated(controls[[j]]$Name), ]
  } 
  # merge controls
  control <- do.call("rbind", controls)
  control <- control %>% group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() 
  control <- control[!duplicated(control$Name), ]

  # Find compounds that are in each sample that are also in control
  # but are much more 'significant' in the sample
  #
  # and also compounds that are not in the controls at all
  samples_keep <- samples
  for(j in 1:length(sample_names)) {
    samp <- samples[[j]] %>% filter(samples[[j]]$Name %in% control$Name)
    cont <- control %>% filter(control$Name %in% samples[[j]]$Name)
   
    # ratio is high enough to keep 
    samp_SN <- unlist(samp[order(samp$Name), "PeakSN"])
    cont_SN <- unlist(cont[order(cont$Name), "PeakSN"])
    contrib1 <- samp %>% filter((samp_SN / cont_SN) > ratio_par)
   
    # also, compounds that are *not* in the controls 
    contrib2 <- samples[[j]] %>% filter(!(samples[[j]]$Name %in% control$Name))
    samples_keep[[j]] <- rbind(contrib1, contrib2)
  }
  names(samples_keep) <- sample_names
  samples_keep
}


##
#
#  join_common: Function which takes output of clean_common above,
#    and merges based on common presence across all replicates of compounds. 
#
#  Inputs:
#  * compounds: list of data.frames, 16 columns as in the spreadsheets
#  
#  Outputs:
#  * common: merged, simplified data.frame, created via inner_join of data.frames after filtering.
##
join_common <- function(compounds) {
  n_samp <- length(compounds)
  subset_compounds <- vector("list", length = n_samp)
  for(j in 1:n_samp) {
    subset_compounds[[j]] <- compounds[[j]]
    if(n_samp > 1) {
      for(k in (1:n_samp)[-j]) {
        subset_compounds[[j]] <- subset_compounds[[j]] %>%
                                   subset(subset_compounds[[j]]$Name %in% compounds[[k]]$Name)
      }
    }
    subset_compounds[[j]] <- subset_compounds[[j]] %>% select(Name, Area, PeakSN)
  }
  
  # Join first two, if they exist
  if(n_samp > 1) {
    common <- inner_join(x = subset_compounds[[1]], y = subset_compounds[[2]], by = "Name")
    if(n_samp >= 3) {
      for(j in 3:n_samp) {
        common <- inner_join(x = common, y = subset_compounds[[j]], by = "Name")  
      }
    }
  } else {
    common <- subset_compounds[[1]][, c("Name", "Area", "PeakSN")]
  }
  names(common) <- c("Name", paste0(c("Area_", "PeakSN_"), rep(1:n_samp, each = 2)))
  common
}
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_1n.rda")
M_DONOR_1 <- join_common( clean_common(M_DONOR_1m,
                         sample_names = c("MR_20H1_1", "MR_20H1_2","MR_20H1_3"),
                         control_names = c("MR_20CH1"),
                         ratio_par = 2.0) )
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_2n.rda")
M_DONOR_2 <- join_common( clean_common(M_DONOR_2m,
                         sample_names = c("MR_20H2_1", "MR_20H2_2","MR_20H2_3"),
                         control_names = c("D2_MControl"),
                         ratio_par = 2.0) )
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_3n.rda")
M_DONOR_3 <- join_common( clean_common(M_DONOR_3m,
                         sample_names = c("MR_20H3_1", "MR_20H3_2","MR_20H3_3"),
                         control_names = c("MR_20CH3"),
                         ratio_par = 2.0) )
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_4n.rda")
M_DONOR_4 <- join_common( clean_common(M_DONOR_4m,
                         sample_names = c("MR_20H4_1", "MR_20H4_2","MR_20H4_3"),
                         control_names = c("MR_20CH4"),
                         ratio_par = 2.0) )
```

```{r, warning = FALSE}
load("./dat/Normalized_RDA/M_DONOR_6n.rda")
M_DONOR_6 <- join_common( clean_common(M_DONOR_6m,
                         sample_names = c("MR_21H6_1", "MR_21H6_2","MR_21H6_3"),
                         control_names = c("MR_21CH6"),
                         ratio_par = 2.0) )
```

```{r, warning = FALSE}
load("./dat/Normalized_RDA/M_DONOR_7n.rda")
M_DONOR_7 <- join_common( clean_common(M_DONOR_7m,
                         sample_names = c("MR_21H7_1", "MR_21H72","MR_21H7_3"),
                         control_names = c("MR_21CH7"),
                         ratio_par = 2.0) )
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_8n.rda")
M_DONOR_8 <- join_common( clean_common(M_DONOR_8m,
                         sample_names = c("MR_21H8_1", "MR_21H8_2","MR_21H8_3"),
                         control_names = c("MR_21CH8"),
                         ratio_par = 2.0) )
```

```{r}
load("./dat/Normalized_RDA/M_DONOR_9n.rda")
M_DONOR_9 <- join_common( clean_common(M_DONOR_9m,
                         sample_names = c("MR_21H9_1", "MR_21H9_2","MR_21H9_3"),
                         control_names = c("MR_21CH9"),
                         ratio_par = 2.0) )
```

## Organize The Content, Write Out

```{r}
all_SS <- vector("list", 9)

for(j in c(1:4, 6:9)) {
  all_SS[[j]] <- get(paste0("M_DONOR_", j))
}
names(all_SS) <- paste0("M_DONOR_", c(1:4, "NA", 6:9))
                   
save(file = "./Morgue_PCA.rda", all_SS)
```

## Cleanup, with Normalization Done First: REST

The REST data, and the cleanup, cannot be done quite as easily, because there
are multiple days being considered for each donor. RH1 is easy, because it was
only recorded for one day; the others, not so much.

```{r}
load("./dat/Normalized_RDA/RH1n.rda")
RH1 <- join_common( clean_common(RH1m,
                                 sample_names = c("D0D11010", "D0D11042","D0D11055"),
                                 control_names = c("RC"),
                                 ratio_par = 2.0) )
```

```{r REST_donor_2}
# Load the normalized donor data
load("./dat/Normalized_RDA/RH2n.rda")
# simple object name
RH2 <- RH2m

# Day 0 - extract, clean, join
RH2_0 <- join_common( clean_common(RH2,
                sample_names = c("D0D21257", "D0D21450","D0D21700"),
                control_names = c("RC"),
                ratio_par = 2.0) )
# Day 1
RH2_1 <- join_common( clean_common(RH2,
                sample_names = c("D1D2945", "D1D21210","D1D21400", "D1D211600"),
                control_names = c("D1D2RC"),
                ratio_par = 2.0) )

# Day 2
RH2_2 <- join_common( clean_common(RH2,
                sample_names = c("D2D21051", "D2D21248","D2D21450", "D2D21650"),
                control_names = c("D2D2RC"),
                ratio_par = 2.0) )

# Day 3
RH2_3 <- join_common( clean_common(RH2,
                sample_names = c("D3D2930", "D3D21130","D3D21328", "D3D21529"),
                control_names = c("D3D2RC"),
                ratio_par = 2.0) )

# Day 4
RH2_4 <- join_common( clean_common(RH2,
                sample_names = c("H2D4R1"),
                control_names = c("D4601934"),
                ratio_par = 2.0) )
```

Continue for the rest of the donors in the same way.

```{r REST_Donor_3, warning = FALSE}
load("./dat/Normalized_RDA/RH3n.rda")
RH3 <- RH3m

# Day 0
RH3_0 <- join_common( clean_common(RH3,
                sample_names = c("D0D3R1", "D0D3R2","D0D3R3"),
                control_names = c("D0D3RC"),
                ratio_par = 2.0) )

# Day 1 Samples were not collected due to the rainfall during the sampling day
RH3_1 <- NULL

# Day 2
RH3_2 <- join_common( clean_common(RH3,
                sample_names = c("D2D3R1", "D2D3R2","D3D3R3"),
                control_names = c("D2D3RC_2"),
                ratio_par = 2.0) )

# Day 3
RH3_3 <- join_common( clean_common(RH3,
                sample_names = c("D3D3R1", "D3D3R2","D3D3R3"),
                control_names = c("D3D3RC"),
                ratio_par = 2.0) )

# Day 4
RH3_4 <- join_common( clean_common(RH3,
                sample_names = c("D4D3R1", "D4D3R2"),
                control_names = c("D4D3RC"),
                ratio_par = 2.0) )
```

And on and on we go ...

```{r REST_Donor_4}
load("./dat/Normalized_RDA/RH4n.rda")
RH4 <- RH4m

# Day 0
RH4_0 <- join_common( clean_common(RH4,
                sample_names = c("D0D4R1", "D0D4R2","D0D4R3"),
                control_names = c("D4RC"),
                ratio_par = 2.0) )

# Day 1
RH4_1 <- join_common( clean_common(RH4,
                sample_names = c("D1D4R1", "D1D4R2","D1D4R3"),
                control_names = c("D1D4RC"),
                ratio_par = 2.0) )

# Day 2 Sampling was not collected due to the rainfall during the sampling day
RH4_2 <- NULL

# Day 3
RH4_3 <- join_common( clean_common(RH4,
                sample_names = c("D3D4R1", "D3D4R2","D3D4R3"),
                control_names = c("D3D4RC"),
                ratio_par = 2.0) )

# Day 4, only control collected
RH4_4 <- NULL
```

```{r REST_Donor_6}
load("./dat/Normalized_RDA/RH6n.rda")
RH6 <- RH6m

# Day 0
RH6_0 <- join_common( clean_common(RH6,
                sample_names = c("D6D0R1", "D6D0R2", "D6D0R3"),
                control_names = c("D6RC"),
                ratio_par = 2.0) )

# Day 1
RH6_1 <- join_common( clean_common(RH6,
                sample_names = c("D6D111_15", "D6D113_10", "D6D114_10"),
                control_names = c("D6D1RC"),
                ratio_par = 2.0) )

# Day 2
RH6_2 <- join_common( clean_common(RH6,
                sample_names = c("D6D211_40", "D6D213_44", "D6D215_30"),
                control_names = c("D6D2RC"),
                ratio_par = 2.0) )

# Day 3
RH6_3 <- join_common( clean_common(RH6,
                sample_names = c("D6D311_41", "D6D313_42", "D6D315_14"),
                control_names = c("D6D3RC"),
                ratio_par = 2.0) )

# Day 4
RH6_4 <- join_common( clean_common(RH6,
                sample_names = c("D6D411_15", "D6D413", "D6D415"),
                control_names = c("D6D4RC"),
                ratio_par = 2.0) )

# Day 5
RH6_5 <- join_common( clean_common(RH6,
                sample_names = c("D6D511_15", "D6D513_45", "D6D515_20"),
                control_names = c("D6D5RC"),
                ratio_par = 2.0) )
```

```{r M_DONOR_7}
load("./dat/Normalized_RDA/RH7n.rda")
RH7 <- RH7m

# Day 0
RH7_0 <- join_common( clean_common(RH7,
                sample_names = c("D0H713_25", "D0H715_30"),
                control_names = c("D1RC"),
                ratio_par = 2.0) )

# Day 1
RH7_1 <- join_common( clean_common(RH7,
                sample_names = c("D1H710_58", "D1H713_21", "D1H715_21"),
                control_names = c("D1RC"),
                ratio_par = 2.0) )

# Day 2
RH7_2 <- join_common( clean_common(RH7,
                sample_names = c("D2H710_45", "D2H713_33", "D2H715_15"),
                control_names = c("D2RC"),
                ratio_par = 2.0) )

# Day 3 Sampling was not collected due to the rainfall during the sampling day
RH7_3 <- NULL

# Day 4
RH7_4 <- join_common( clean_common(RH7,
                sample_names = c("D4H711_30", "D4H713_40", "D4H715_10"),
                control_names = c("D4RC"),
                ratio_par = 2.0) )
# Day 5
RH7_5 <- join_common( clean_common(RH7,
                sample_names = c("D5H79_45", "D5H712_45", "D5H714_45"),
                control_names = c("D5RC"),
                ratio_par = 2.0) )
```

```{r M_DONOR_8}
load("./dat/Normalized_RDA/RH8n.rda")
RH8 <- RH8m

# Day 0
RH8_0 <- join_common( clean_common(RH8,
                sample_names = c("H8D0R1","H8D0R2"),
                control_names = c("H8D0RC"),
                ratio_par = 2.0) )

# Day 1
RH8_1 <- join_common( clean_common(RH8,
                sample_names = c("H8D1R1","H8D1R2", "H8D1R3"),
                control_names = c("H8D1RC"),
                ratio_par = 2.0) )
# Day 2
RH8_2 <- join_common( clean_common(RH8,
                sample_names = c("H8D2R1","H8D2R2", "H8D2R3"),
                control_names = c("D2H8RC"),
                ratio_par = 2.0) )
# Day 3
RH8_3 <- join_common( clean_common(RH8,
                sample_names = c("H8D3R1","H8D3R2", "H8D3R3"),
                control_names = c("H8D3RC"),
                ratio_par = 2.0) )
# Day 4  
RH8_4 <- join_common( clean_common(RH8,
#                sample_names = c("RD_RC_Tube_3_AM_sample", "H8D4R3"),  #### NOTE: nothing survives from RD_RC
                sample_names = "H8D4R3",
                control_names = c("H8D4RC"),
                ratio_par = 2.0) )

# Day 5
RH8_5 <- join_common( clean_common(RH8,
                sample_names = c("H8D5R1"),
                control_names = c("H8D5RC"),
                ratio_par = 2.0) )
```

```{r REST Donor 9}
load("./dat/Normalized_RDA/RH9n.rda")
RH9 <- RH9m

# Day 0
RH9_0 <- join_common( clean_common(RH9,
                sample_names = c("D0H910_55", "D0H913_18", "D0H915_17"),
                control_names = c("D0RC"),
                ratio_par = 2.0) )

# Day 1
RH9_1 <- join_common( clean_common(RH9,
                sample_names = c("D1H910_36", "D1H913_12", "D1H915_34"),
                control_names = c("D1RC (2)"),
                ratio_par = 2.0) )

# Day 2
RH9_2 <- join_common( clean_common(RH9,
                sample_names = c("D2H910_53", "D2RC13_32", "D2RC15_11"),
                control_names = c("D2RC"),
                ratio_par = 2.0) )

# Day 3
RH9_3 <- join_common( clean_common(RH9,
                sample_names = c("D3H911_1", "D3H913_25", "D3H915_21"),
                control_names = c("D3RC"),
                ratio_par = 2.0) )

# Day 4
RH9_4 <- join_common( clean_common(RH9,
                sample_names = c("D4H910_50", "D4H913_18", "D4H915_35"),
                control_names = c("D4RC"),  
                ratio_par = 2.0) )

# Day 5
RH9_5 <- join_common( clean_common(RH9,
                sample_names = c("D5H910:58"),
                control_names = c("D5RC"),
                ratio_par = 2.0) )
```

## Organize The Content, Write Out

```{r}
all_SS <- vector("list", 35)  # total number of donor+day

# DONOR 1
all_SS[[1]] <- RH1
# DONOR 2
all_SS[[2]] <- RH2_0
all_SS[[3]] <- RH2_1
all_SS[[4]] <- RH2_2
all_SS[[5]] <- RH2_3 
all_SS[[6]] <- RH2_4
# DONOR 3
all_SS[[7]]  <- RH3_0
all_SS[[8]]  <- RH3_2
all_SS[[9]]  <- RH3_3 
all_SS[[10]] <- RH3_4
# DONOR 4
all_SS[[11]]  <- RH4_0
all_SS[[12]]  <- RH4_1
all_SS[[13]]  <- RH4_3 
# DONOR 6
all_SS[[14]]  <- RH6_0
all_SS[[15]]  <- RH6_1
all_SS[[16]]  <- RH6_2 
all_SS[[17]]  <- RH6_3
all_SS[[18]]  <- RH6_4
all_SS[[19]]  <- RH6_5 
# DONOR 7
all_SS[[20]]  <- RH7_0
all_SS[[21]]  <- RH7_1 
all_SS[[22]]  <- RH7_2
all_SS[[23]]  <- RH7_4
all_SS[[24]]  <- RH7_5 
# DONOR 8
all_SS[[25]]  <- RH8_0
all_SS[[26]]  <- RH8_1
all_SS[[27]]  <- RH8_2
all_SS[[28]]  <- RH8_3
all_SS[[29]]  <- RH8_4
all_SS[[30]]  <- RH8_5
# DONOR 9
all_SS[[31]]  <- RH9_0
all_SS[[32]]  <- RH9_1
all_SS[[33]]  <- RH9_2
all_SS[[34]]  <- RH9_3
all_SS[[35]]  <- RH9_4
all_SS[[36]]  <- RH9_5

names(all_SS) <- c("D1", 
                   "D2_d0", "D2_d1", "D2_d2", "D2_d3", "D2_d4",
                   "D3_d0", "D3_d2", "D3_d3", "D3_d4",
                   "D4_d0", "D4_d1", "D4_d3",
                   "D6_d0", "D6_d1", "D6_d2", "D6_d3", "D6_d4", "D6_d5",
                   "D7_d0", "D7_d1", "D7_d2", "D7_d4", "D7_d5",
                   "D8_d0", "D8_d1", "D8_d2", "D8_d3", "D8_d4", "D8_d5",
                   "D9_d0", "D9_d1", "D9_d2", "D9_d3", "D9_d4", "D9_d5")

save(file = "./REST_PCA.rda", all_SS)
```

