---
title: "Filter Extraneous Compounds"
author: "Wesley Burr"
date: "`r Sys.Date()`"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## REST: Cleanup Noted Compounds Throughout

There are a number of compounds that have to do with the
column, the reference standard (somehow coming through), 
and other definitely not-related things. We will strip
them out now, in preparation for PCA.

### Remove Explicitly

The following compounds are definitely not related to decomp,
and are related to the process or environment, and need to be
removed:

* Bromobenzene (or Benzene, bromo) 
* Oxygen
* Acetone
* Methyl alcohol / Methanol
* Carbon dioxide

Before we start that, we'll make a list of the actual
"appearances" of compounds, and then check against this
to determine the actual filtering arguments (e.g., first 8 characters
or full name, etc.).

```{r}
library(tidyverse)
library(readxl)
load("./REST_PCA.rda")
all_compounds <- sort(unique(
  unlist(lapply(all_SS, FUN = function(x) { unlist(x$Name) }))))
length(all_compounds)
```

So there are `r length(all_compounds)` (used to be 1438) 
unique compounds present across the 35
total samples, after Controlling (but with the 2.0 ratio
argument in place). Let's look for each of the above compounds:

```{r benz}
loc1 <- grep(pattern = "Benzene, bromo-", all_compounds, 
             ignore.case = TRUE)
loc2 <- grep(pattern = "oxygen", all_compounds, ignore.case = TRUE)
loc3 <- grep(pattern = "acetone", all_compounds, ignore.case = TRUE)
loc4 <- grep(pattern = "methyl alc", all_compounds, ignore.case = TRUE)
loc5 <- grep(pattern = "methanol, TMS", all_compounds, 
             ignore.case = TRUE)
loc6 <- grep(pattern = "carbon diox", all_compounds, ignore.case = TRUE)
```

Put them together as indexes, then extract these from the
list of compounds as actual names.

```{r}
remove_specifics <- c(loc1, loc2, loc3, loc4, loc5, loc6)
remove1 <- all_compounds[remove_specifics]
```

### Remove Via Keyword

There are three keywords that show up that we should also
strip out:

* Sil
* TMS
* TBDMS

Let's grab these now:

```{r}
loc1 <- grep(pattern = "Sil", all_compounds, ignore.case = TRUE)
loc2 <- grep(pattern = "TMS", all_compounds)
loc3 <- grep(pattern = "TBDMS", all_compounds)
remove2 <- all_compounds[c(loc1, loc2, loc3)]
```

### Merge Compounds

We need some logic to look for things that are the same
compound, but different in only stereochemistry. The
indicator seems to be brackets: (E), (Z), (S),
popping up in one or more of the variants. So there might be,
for example, 2-Octene, (E)- as a compound, and then another
sample might have 2-Octene, (Z)-. These should just be merged. 

Let's try to look for them first:

```{r}
loc1 <- grep(pattern = "\\(E\\)-$", all_compounds)
loc2 <- grep(pattern = "\\(Z\\)-$", all_compounds)
loc3 <- grep(pattern = "\\(S\\)-$", all_compounds)
loc4 <- grep(pattern = "\\(R\\)-$", all_compounds)
to_clean <- c(loc1, loc2, loc3, loc4)
```

Now, the tricky bit: how to fix this up. What we want
are these compounds, and their corresponding compounds which
**don't** have the (S), (Z), (R) or (E); or have a different one.
In all cases, we'll merge them down to the **doesn't have brackets**
version if it exists, or if it doesn't, we'll make one.

```{r}
mappings <- data.frame(Original = NA, Transformed = NA)
for(j in 1:length(to_clean)) {
  orig <- all_compounds[to_clean[j]]
  fixed <- strsplit(orig, "\\(")[[1]][1]
  fixed <- substr(fixed, 1, nchar(fixed) - 2)
  mappings[j, ] <- c(orig, 
                     fixed)
}
```

## Back to the REST Data, Ready to Rock & Roll

So we have remove1 - compounds to remove. We have remove2 - 
more compounds to remove. And we have mappings, which have
compounds that need to be renamed. Then, at the end, we need
to check for duplicates, because the renaming may have 
resulted in more than one compound surviving in a single sample
due to the stereochemistry issue. 

```{r}
test <- lapply(all_SS, FUN = function(x) { 
    y <- x %>% filter(!(Name %in% remove1 | Name %in% remove2))
    which_rows <- which(y$Name %in% mappings$Original)
    if(length(which_rows) > 0) {
      for(j in 1:length(which_rows)) {
        orig <- unlist(y[which_rows[j], "Name"])
        y[which_rows[j], "Name"] <- mappings[mappings$Original == orig,
                                             "Transformed"]
      }
    }
    y
  })
```

Now, look for duplicates, and remove if any now exist:

```{r}
test <- lapply(test, FUN = function(x) { 
    dupes <- which(duplicated(x$Name))
    if(length(dupes) > 0) {
      x[-dupes, ]
    } else {
      x
    }
  })
```

We're done! Ready to send this off to PCA using other software.

```{r}
library("xlsx")
write.xlsx(x = test[[1]],
           file = "./all_REST_cleaned.xlsx",
           sheetName = names(test)[1],
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 2:length(test)) {
  write.xlsx(x = test[[j]],
             file = "./all_REST_cleaned.xlsx",
             sheetName = names(test)[j],
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE)
}
```

## Extract Top Compounds

```{r}
compounds <- unlist(lapply(test, FUN = function(x) { x$Name }))
summary_table <- table(compounds)
summary_table <- sort(summary_table, decreasing = TRUE)
top_compounds <- summary_table[1:(min(which(summary_table < 2))-1)]
output_df <- data.frame(top_compounds)
names(output_df) <- c("Compound", "Frequency (out of 36)")
write.csv(file = "./REST_top_compounds.csv", output_df,
          row.names = FALSE)
```

We also want to know which samples they appeared in, which is
less easy to do. We have the compound names at least, so now
we need to create 36 variables to track things, and then fill it in.

```{r}
output_df2 <- as.data.frame(matrix(data = NA, nrow = nrow(output_df),
                                   ncol = 36))
names(output_df2) <- names(test)
output_df2 <- cbind(output_df, output_df2)
for(j in 1:nrow(output_df)) {
    comp <- output_df[j, "Compound"]
    fit <- unlist(lapply(test, FUN = function(x) { 
      if(comp %in% x$Name) { 1 } else { 0 }}))
    output_df2[j, -(1:2)] <- fit
}
write.csv(file = "./REST_top_compounds_withSpecifics.csv", output_df2,
          row.names = FALSE)
```

## Morgue: Cleanup Noted Compounds Throughout

There are a number of compounds that have to do with the
column, the reference standard (somehow coming through), 
and other definitely not-related things. We will strip
them out now, in preparation for PCA. This is identical
to the process for the REST data, but done on a different
source file (and different numbering).

### Remove Explicitly

The following compounds are definitely not related to decomp,
and are related to the process or environment, and need to be
removed:

* Bromobenzene (or Benzene, bromo) 
* Oxygen
* Acetone
* Methyl alcohol / Methanol
* Carbon dioxide

Before we start that, we'll make a list of the actual
"appearances" of compounds, and then check against this
to determine the actual filtering arguments (e.g., first 8 characters
or full name, etc.).

```{r}
library(tidyverse)
library(readxl)
load("./Morgue_PCA.rda")
all_compounds <- sort(unique(
  unlist(lapply(all_SS, FUN = function(x) { unlist(x$Name) }))))
length(all_compounds)
```

So there are `r length(all_compounds)` 
unique compounds present across the 9 
total samples, after Controlling (but with the 2.0 ratio
argument in place). Let's look for each of the above compounds:

```{r benz2}
loc1 <- grep(pattern = "Benzene, bromo-", all_compounds, 
             ignore.case = TRUE)
loc2 <- grep(pattern = "oxygen", all_compounds, ignore.case = TRUE)
loc3 <- grep(pattern = "acetone", all_compounds, ignore.case = TRUE)
loc4 <- grep(pattern = "methyl alc", all_compounds, ignore.case = TRUE)
loc5 <- grep(pattern = "methanol, TMS", all_compounds, 
             ignore.case = TRUE)
loc6 <- grep(pattern = "carbon diox", all_compounds, ignore.case = TRUE)
```

Put them together as indexes, then extract these from the
list of compounds as actual names.

```{r}
remove_specifics <- c(loc1, loc2, loc3, loc4, loc5, loc6)
remove1 <- all_compounds[remove_specifics]
```

### Remove Via Keyword

There are three keywords that show up that we should also
strip out:

* Sil
* TMS
* TBDMS

Let's grab these now:

```{r}
loc1 <- grep(pattern = "Sil", all_compounds, ignore.case = TRUE)
loc2 <- grep(pattern = "TMS", all_compounds)
loc3 <- grep(pattern = "TBDMS", all_compounds)
remove2 <- all_compounds[c(loc1, loc2, loc3)]
```

### Merge Compounds

We need some logic to look for things that are the same
compound, but different in only stereochemistry. The
indicator seems to be brackets: (E), (Z), (S),
popping up in one or more of the variants. So there might be,
for example, 2-Octene, (E)- as a compound, and then another
sample might have 2-Octene, (Z)-. These should just be merged. 

Let's try to look for them first:

```{r}
loc1 <- grep(pattern = "\\(E\\)-$", all_compounds)
loc2 <- grep(pattern = "\\(Z\\)-$", all_compounds)
loc3 <- grep(pattern = "\\(S\\)-$", all_compounds)
loc4 <- grep(pattern = "\\(R\\)-$", all_compounds)
to_clean <- c(loc1, loc2, loc3, loc4)
```

Now, the tricky bit: how to fix this up. What we want
are these compounds, and their corresponding compounds which
**don't** have the (S), (Z), (R) or (E); or have a different one.
In all cases, we'll merge them down to the **doesn't have brackets**
version if it exists, or if it doesn't, we'll make one.

```{r}
mappings <- data.frame(Original = NA, Transformed = NA)
for(j in 1:length(to_clean)) {
  orig <- all_compounds[to_clean[j]]
  fixed <- strsplit(orig, "\\(")[[1]][1]
  fixed <- substr(fixed, 1, nchar(fixed) - 2)
  mappings[j, ] <- c(orig, 
                     fixed)
}
```

## Back to the Morgue Data, Ready to Rock & Roll

So we have remove1 - compounds to remove. We have remove2 - 
more compounds to remove. And we have mappings, which have
compounds that need to be renamed. Then, at the end, we need
to check for duplicates, because the renaming may have 
resulted in more than one compound surviving in a single sample
due to the stereochemistry issue. 

```{r}
test <- lapply(all_SS[-5], FUN = function(x) { 
    y <- x %>% filter(!(Name %in% remove1 | Name %in% remove2))
    which_rows <- which(y$Name %in% mappings$Original)
    if(length(which_rows) > 0) {
      for(j in 1:length(which_rows)) {
        orig <- unlist(y[which_rows[j], "Name"])
        y[which_rows[j], "Name"] <- mappings[mappings$Original == orig,
                                             "Transformed"]
      }
    }
    y
  })
```

Now, look for duplicates, and remove if any now exist:

```{r}
test <- lapply(test, FUN = function(x) { 
    dupes <- which(duplicated(x$Name))
    if(length(dupes) > 0) {
      x[-dupes, ]
    } else {
      x
    }
  })
```

We're done! Ready to send this off to PCA using other software.

```{r}
library("xlsx")
write.xlsx(x = test[[1]],
           file = "./all_Morgue_cleaned.xlsx",
           sheetName = names(test)[1],
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 2:length(test)) {
  write.xlsx(x = test[[j]],
             file = "./all_Morgue_cleaned.xlsx",
             sheetName = names(test)[j],
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE)
}
```

## Extract Top Compounds

```{r}
compounds <- unlist(lapply(test, FUN = function(x) { x$Name }))
summary_table <- table(compounds)
summary_table <- sort(summary_table, decreasing = TRUE)
top_compounds <- summary_table[1:(min(which(summary_table < 2))-1)]
output_df <- data.frame(top_compounds)
names(output_df) <- c("Compound", "Frequency (out of 8)")
write.csv(file = "./Morgue_top_compounds.csv", output_df,
          row.names = FALSE)
```

We also want to know which samples they appeared in, which is
less easy to do. We have the compound names at least, so now
we need to create 8 variables to track things, and then fill it in.

```{r}
output_df2 <- as.data.frame(matrix(data = NA, nrow = nrow(output_df),
                                   ncol = 8))
names(output_df2) <- names(test)
output_df2 <- cbind(output_df, output_df2)
for(j in 1:nrow(output_df)) {
    comp <- output_df[j, "Compound"]
    fit <- unlist(lapply(test, FUN = function(x) { 
      if(comp %in% x$Name) { 1 } else { 0 }}))
    output_df2[j, -(1:2)] <- fit
}
write.csv(file = "./Morgue_top_compounds_withSpecifics.csv", output_df2,
          row.names = FALSE)
```

