---
title: "Examine Benzene RTs"
author: "Wesley Burr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file requires that the 1_Extract_Setup.Rmd and 1_UN_Extract_Setup.Rmd
files have been run. This sets up the rda files in dat and dat_un so this
can work. 

## Extract the RTs for the Max SN Bromobenzene (Aligned)

```{r}
library("tidyverse")
bromo_Morgue <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat/M_DONOR_", k, ".rda"))
  M_DONOR_ <- get(x = paste0("M_DONOR_", k))
  bromo <- M_DONOR_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(QuantSN == max(QuantSN))
  bromo_Morgue[[k]] <- bromo 
}
names(bromo_Morgue) <- paste0("DONOR_", 1:9)
```


```{r}
bromo_REST <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat/RH", k, ".rda"))
  RH_ <- get(x = paste0("RH", k))
  bromo <- RH_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(QuantSN == max(QuantSN))
  bromo_REST[[k]] <- bromo
}
names(bromo_REST) <- paste0("RH_", 1:9)
```

## Cleanup, Make Excel

```{r}
s1 <- do.call("rbind", bromo_Morgue)
s1 <- s1 %>% select(Sample, RT1, RT2)
s2 <- do.call("rbind", bromo_REST)
s2 <- s2 %>% select(Sample, RT1, RT2)
bromo <- rbind(s1, s2)
bromo <- t(bromo)
bromo <- data.frame(bromo)
names(bromo) <- bromo[1, ]
bromo <- bromo[-1, ]
bromo <- apply(bromo, MAR = 2, FUN = as.numeric)
bromo <- as.data.frame(bromo)
bromo
```

## Write Out File

```{r}
write.csv(file = "Checking_Bromobenze_IS.csv", bromo, row.names = FALSE)
```


```{r}
plot(x = unlist(bromo[1, ]),
     y = unlist(bromo[2, ]),
     xlab = "RT1", 
     ylab = "RT2")
```

## Extract the RTs for the Max SN Bromobenzene (Unaligned)

```{r}
bromo_Morgue <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat_un/M_DONOR_", k, ".rda"))
  M_DONOR_ <- get(x = paste0("M_DONOR_", k))
  bromo <- M_DONOR_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(QuantSN == max(QuantSN))
  bromo_Morgue[[k]] <- bromo 
}
names(bromo_Morgue) <- paste0("DONOR_", 1:9)
```


```{r}
bromo_REST <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat_un/RH", k, ".rda"))
  RH_ <- get(x = paste0("RH", k))
  bromo <- RH_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(QuantSN == max(QuantSN))
  bromo_REST[[k]] <- bromo
}
names(bromo_REST) <- paste0("RH_", 1:9)
```

## Cleanup, Make Excel

```{r}
s1 <- do.call("rbind", bromo_Morgue)
s1 <- s1 %>% select(Sample, RT1, RT2)
s2 <- do.call("rbind", bromo_REST)
s2 <- s2 %>% select(Sample, RT1, RT2)
bromo <- rbind(s1, s2)
bromo <- t(bromo)
bromo <- data.frame(bromo)
names(bromo) <- bromo[1, ]
bromo <- bromo[-1, ]
bromo <- apply(bromo, MAR = 2, FUN = as.numeric)
bromo <- as.data.frame(bromo)
bromo
```

## Write Out File

```{r}
write.csv(file = "Checking_Bromobenze_IS_un.csv", bromo, row.names = FALSE)
```


```{r}
plot(x = unlist(bromo[1, ]),
     y = unlist(bromo[2, ]),
     xlab = "RT1", 
     ylab = "RT2")
```

## Extract, Find Medians in Both Dimensions - from Aligned



```{r}
bromo_Morgue <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat/M_DONOR_", k, ".rda"))
  M_DONOR_ <- get(x = paste0("M_DONOR_", k))
  bromo <- M_DONOR_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           top_n(wt = PeakSN, n = 2) %>%
           group_split()
  bromo_Morgue[[k]] <- do.call("rbind", bromo)
}
names(bromo_Morgue) <- paste0("DONOR_", 1:9)

bromo_REST <- vector("list", length = 9)
for(k in c(1:4, 6:9)) {
  load(paste0("./dat/RH", k, ".rda"))
  RH_ <- get(x = paste0("RH", k))
  bromo <- RH_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           top_n(wt = PeakSN, n = 2) %>%
           group_split()
  bromo_REST[[k]] <- do.call("rbind", bromo)
}
names(bromo_REST) <- paste0("RH_", 1:9)

bromo <- rbind(do.call("rbind", bromo_Morgue) %>% select(RT1, RT2, BaseMass, PeakSN), 
               do.call("rbind", bromo_REST) %>% select(RT1, RT2, BaseMass, PeakSN))
```

From this, we find that we have `r nrow(bromo)` recorded records of Bromobenzene
in our data set. The median RTs are `r median(bromo$RT1)` and `r median(bromo$RT2)`.
Plotted, they look like:

```{r}
cols <- rep("black", nrow(bromo))
cols[bromo$BaseMass <= 77.10 & bromo$BaseMass >= 76.90] <- "red"
plot(bromo$RT1, bromo$RT2, col = cols)
```

```{r}
sliding_window <- function(dat, xww, yww) {

  # dat has RT1, RT2 and BaseMass in it - find the most common mass
  filt <- hist(dat$BaseMass, breaks = 1:(max(dat$BaseMass) + 1),
               plot = FALSE)
  common_mass <- filt$mids[filt$counts == max(filt$counts)]
  dat <- dat %>% filter(BaseMass <= common_mass * 1.02, 
                        BaseMass >= common_mass * 0.98)
  
  min_x <- min(dat$RT1)
  max_x <- max(dat$RT1)
  min_y <- min(dat$RT2)
  max_y <- max(dat$RT2)
  
  # Form a set of bottom-left to top-right boundaries, coarse - half the 
  # widths xww and yww as steps
  x_seq <- seq(from = min_x - xww, to = max_x + xww, by = xww / 2)
  y_seq <- seq(from = min_y - yww, to = max_y + yww, length.out = length(x_seq))
  boundaries1 <- data.frame(x1 = rep(x_seq, each = length(y_seq)), 
                            x2 = rep(x_seq + xww, each = length(y_seq)),
                            y1 = rep(y_seq, length(x_seq)),
                            y2 = rep(y_seq + yww, length(x_seq)))
  first_pass <- apply(boundaries1, MAR = 1, FUN = function(z) {
                        nrow(dat[dat$RT1 >= z[1] & dat$RT1 <= z[2] & 
                             dat$RT2 >= z[3] & dat$RT2 <= z[4], ])
                    })
  
  # identify the area of local maximum, then zoom in there and redo this
  # with a much finer x-resolution (that's the limiting factor)
  which_max <- which(first_pass == max(first_pass))[1]
  x_seq <- seq(from = boundaries1[which_max, "x1"] - 2 * xww,
               to = boundaries1[which_max, "x2"] + xww,
               by = xww / 10)
  y_seq <- seq(from = boundaries1[which_max, "y1"] - 2 * yww,
               to = boundaries1[which_max, "y2"] + yww,
               by = yww / 10)
  boundaries2 <- data.frame(x1 = rep(x_seq, each = length(y_seq)), 
                            x2 = rep(x_seq + xww, each = length(y_seq)),
                            y1 = rep(y_seq, length(x_seq)),
                            y2 = rep(y_seq + yww, length(x_seq)))
  
  second_pass <- apply(boundaries2, MAR = 1, FUN = function(z) {
                         nrow(dat[dat$RT1 >= z[1] & dat$RT1 <= z[2] & 
                              dat$RT2 >= z[3] & dat$RT2 <= z[4], ])
                     })
  x <- mean(unlist(boundaries2[which(second_pass == max(second_pass))[1], 1:2]))
  y <- mean(unlist(boundaries2[which(second_pass == max(second_pass))[1], 3:4]))
  return(c(common_mass, x, y))
}
```

```{r testing}
sliding_window(bromo, 10, 0.05)
sliding_window(bromo, 10, 0.10)
sliding_window(bromo, 15, 0.05)
sliding_window(bromo, 15, 0.10)
```

This is the estimated 'centroid' for the common mass (+/-2%),
and the centroid RT1 and RT2 estimates, which match your chosen RT1 and RT2
variability.

We can now apply this method to any compound we like, to find its "definitive"
window, which can then be used for filtering! Time to write some more code ...