---
title: "After Alignment Load & Review + Filtered Bromo"
author: "Wesley"
date: "2022/11/20"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Adding the Filter for RT and Mass

In this iteration, we use the filtering for RT and common mass that we derived
in the 7_Identifying_Commons file. This is a rda.

```{r}
load("compound_ref_RT_mass.rda")
```

The format of this object is a data.frame, with the Name, common_mass,
window_RT1 center and window_RT2 centers. This will let us filter down our
data to only those compounds matching these requirements, so we can be sure
that whatever we find at the end is well-matched across GCxGC samples and
time. 

```{r}
for(k in c(1:4, 6:9)) {
  load(paste0("./dat/M_DONOR_", k, ".rda"))
  load(paste0("./dat/RH", k, ".rda"))
}

RT1_ww <- 5.0
RT2_ww <- 0.1
```

## Normalize the Morgue Data Against Their Internal Standards

```{r}
if(!dir.exists("./dat/Normalized_RDA")) {
  dir.create("./dat/Normalized_RDA")
}

ref_bromo <- data.frame(Name = "Benzene, bromo-", 
                        common_mass = 77.50, 
                        window_RT1 = 1114.93, 
                        window_RT2 = 0.79)

for(k in c(1:4, 6:9)) {
  M_DONOR_ <- get(x = paste0("M_DONOR_", k))
  bromo <- M_DONOR_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(BaseMass >= 0.98 * ref_bromo$common_mass,  # 'correct' mass
                  BaseMass <= 1.02 * ref_bromo$common_mass,
                  RT1 <= ref_bromo$window_RT1 + RT1_ww,      # 'correct' RTs
                  RT1 >= ref_bromo$window_RT1 - RT1_ww,
                  RT2 <= ref_bromo$window_RT2 + RT2_ww,
                  RT2 >= ref_bromo$window_RT2 - RT2_ww) %>%
           filter(Area == max(Area)) %>%
           slice_head(n = 1)
  split_M_DONOR_ <- M_DONOR_ %>% 
                group_by(Sample) %>%
                group_split()
  for(j in 1:length(split_M_DONOR_)) {
    split_M_DONOR_[[j]]$Area <- split_M_DONOR_[[j]]$Area / unlist(bromo[j, "Area"])
  }
  M_DONOR_ <- do.call("rbind", split_M_DONOR_)
  assign(x = paste0("M_DONOR_", k, "m"), value = M_DONOR_)
  save(file = paste0("./dat/Normalized_RDA/M_DONOR_", k, "m8.rda"), 
       list = paste0("M_DONOR_", k, "m"))
}
```


## Normalize the REST Data Against Their Internal Standards

```{r}
if(!dir.exists("./dat/Normalized_RDA")) {
  dir.create("./dat/Normalized_RDA")
}

for(k in c(1:4, 6:9)) {
  RH_ <- get(x = paste0("RH", k))
  bromo <- RH_ %>% 
           group_by(Sample) %>% 
           subset(Name == "Benzene, bromo-") %>%
           filter(BaseMass >= 0.98 * ref_bromo$common_mass,  # 'correct' mass
                  BaseMass <= 1.02 * ref_bromo$common_mass,
                  RT1 <= ref_bromo$window_RT1 + RT1_ww,      # 'correct' RTs
                  RT1 >= ref_bromo$window_RT1 - RT1_ww,
                  RT2 <= ref_bromo$window_RT2 + RT2_ww,
                  RT2 >= ref_bromo$window_RT2 - RT2_ww) %>%
           filter(Area == max(Area)) %>%
           slice_head(n = 1)
  split_RH_ <- RH_ %>% 
                group_by(Sample) %>%
                group_split()
  for(j in 1:length(split_RH_)) {
    split_RH_[[j]]$Area <- split_RH_[[j]]$Area / unlist(bromo[j, "Area"])
  }
  RH_ <- do.call("rbind", split_RH_)
  assign(x = paste0("RH", k, "m"), value = RH_)
  save(file = paste0("./dat/Normalized_RDA/RH", k, "m8.rda"), 
       list = paste0("RH", k, "m"))
}
```
